On the number of tiles visited by a\\ line segment on a rectangular grid

### 2021-11-25 paper.tex
Rewriting

\affil[1]{\small Universidad Polit\'ecnica de Madrid. \texttt{luis.mendo@upm.es}}
%Information Processing and Telecommunications Center, Universidad Polit\'ecnica de Madrid\\
%Avenida Complutense, 30. 28040 Madrid, Spain.\\
\affil[2]{\small **Affiliation, e-mail**}

@@

For a grid with horizontal spacing $a$ and vertical spacing $b$, lines of the form $x = ka$ or $y = kb$ with $k \in \mathbb Z$ will be called \emph{grid lines}. A \emph{tile} is delimited by two pairs of consecutive horizontal and vertical grid lines. The intersection points of horizontal and vertical grid lines will be called \emph{grid points}. These correspond to vertices of the tiles.

Every segment has an associated \emph{discrete bounding rectangle}, which is the minimum-size rectangle that is formed by grid lines and contains the segment. More specifically, if the segment has endpoints $(x_1,y_1)$, $(x_2,y_2)$, where $x_1, x_2, y_1, y_2 \in \mathbb R$, its discrete bounding rectangle has lower-left corner and upper-right corner given as
\begin{align*}
& (\lfloor\min\{x_1, x_2\}/a\rfloor a, \lfloor\min\{y_1,y_2\}/b\rfloor b), \\
& (\lceil\max\{x_1, x_2\}/a \rceil a, \lceil\max\{y_1,y_2\}/b \rceil b).
\end{align*}
The dimensions of the discrete bounding rectangle, normalized to the tile width and height respectively, are two integer numbers $i$, $j$. Two examples are illustrated in Figure~\ref{fig: discrete bounding rectangle and touched tiles}, both with $i=5$, $j=4$. Clearly, all tiles visited by the segment are contained in the discrete bounding rectangle. Note also that the rectangle can have $i=0$ or $j=0$ if the segment coincides with part of a grid line.

@@

\begin{proof}
\ref*{prop: len ineq i j: ineqs} The inequalities follow from the fact that the segment endpoints lie in the interiors or on the outer edges of two tiles in opposite corners of the discrete bounding rectangle. This is illustrated in Figures~\ref{fig: len ineq 4 3} and \ref{fig: len ineq 4 2} for two specific $(i,j)$ pairs respectively. For each $(i,j)$, segments are shown with lengths close to either of the two bounds. Note that \eqref{eq: len > sqrt} is valid even for $i=2$, $j=2$, in which case it reduces to $\len>0$.

\ref*{prop: len ineq i j: exist} For $a$, $b$, $\len$, $i$, $j$ satisfying the two inequalities, a segment of length $\len$ can be found that has its endpoints in the interiors or on the outer edges of the two shaded tiles of a rectangle of normalized dimensions $i$ and $j$, as in Figure~\ref{fig: len ineq i j}, which is thus the discrete bounding rectangle of that segment.
\begin{figure}
\centering%
\subfigure[$i=4$, $j=3$]{%
\label{fig: len ineq 4 3}%
\includegraphics[width=.49\textwidth]{L_ineq_4_3_1p35}%
}\hfill%
\subfigure[$i=4$, $j=2$]{%
\label{fig: len ineq 4 2}%
\includegraphics[width=.49\textwidth]{L_ineq_4_2_1p35}%
}%
\caption{Relationship between segment length $\len$ and dimensions $i$, $j$ of the discrete bounding rectangle
}%
\label{fig: len ineq i j}
\end{figure}%

@@

``There is a segment\ldots $\Rightarrow$ \eqref{eq: len > sqrt} '': Assume that a segment exists with length not exceeding $\len$ and with a discrete bounding rectangle with normalized dimensions $i$, $j$. From part~\ref{prop: len ineq i j: ineqs}, inequality \eqref{eq: len > sqrt} holds for that segment length, and thus for $\len$.
\end{proof}

According to Proposition~\ref{prop: i+j-1}, in order to maximize the number of visited tiles for a given length, the position and orientation of the segment should be chosen to obtain $i+j-1$ as large as possible, where $i$ and $j$ are the normalized dimensions of its discrete bounding rectangle. On the other hand, Proposition~\ref{prop: len ineq i j} restricts the $i, j$ values that can be achieved with a given length. A relevant question is: are there any $(i,j)$ pairs that can be disregarded irrespective of the length $\len$? In other words, what is the ``smallest'' subset of $\mathbb N^2$ such that the $(i,j)$ pair that maximizes the number of tiles for any given length can always be found within that subset?

@@

Consider $a=b=1$ first. This is illustrated in Figure~\ref{fig: ijLS_1}. Note that in this and in the next figures the axes represent $ia$ and $jb$, rather than
% normalized coordinates
$i$ and $j$. In this graph, each dashed diagonal line contains pairs $(ia,jb)$ with the same $i+j-1$; and the lower bound \eqref{eq: len > sqrt} corresponds to an arc centered at $(2a,2b)$.

@@

\begin{proof}
Suppose first that the grid is simplified to only horizontal lines with spacing $b$. This matches the setup of Buffon's original needle problem \cite{Mathai99}, except that here the length $\len$ of the needle may exceed the spacing $b$, allowing it to cross multiple lines. As shown in \cite{Ramaley69}, the expected number of lines crossed equals $2\len/(\pi b)$.
% When $\len<b$ this expectation matches the probability of the needle crossing a grid line, since it cannot cross more than one.

>>>>

\affil[1]{\small Universidad Polit\'ecnica de Madrid. \texttt{luis.mendo@upm.es}}
%Information Processing and Telecommunications Center, Universidad Polit\'ecnica de Madrid\\
%Avenida Complutense, 30. 28040 Madrid, Spain.\\
\affil[2]{\small \texttt{arkhipov@alum.mit.edu}}

@@

For a grid with horizontal spacing $a$ and vertical spacing $b$, lines $x = ka$ or $y = kb$ with $k \in \mathbb Z$ will be called \emph{grid lines}. A \emph{tile} is delimited by two pairs of consecutive horizontal and vertical grid lines. The intersection points of horizontal and vertical grid lines will be called \emph{grid points}. These correspond to vertices of the tiles.

Every segment has an associated \emph{discrete bounding rectangle}, which is the minimum-size rectangle that is formed by grid lines and contains the segment. More specifically, if the segment has endpoints $(x_1,y_1)$, $(x_2,y_2) \in \mathbb R^2$, its discrete bounding rectangle has lower-left and upper-right corners respectively given as
\begin{align*}
& (\lfloor\min\{x_1, x_2\}/a\rfloor a, \lfloor\min\{y_1,y_2\}/b\rfloor b), \\
& (\lceil\max\{x_1, x_2\}/a \rceil a, \lceil\max\{y_1,y_2\}/b \rceil b).
\end{align*}
The dimensions of the discrete bounding rectangle, normalized to the tile width and height respectively, are two integer numbers $i$, $j$. Two examples are illustrated in Figure~\ref{fig: discrete bounding rectangle and touched tiles}, both with $i=5$, $j=4$. All tiles visited by the segment are contained in the discrete bounding rectangle. Note also that the rectangle can have $i=0$ or $j=0$ if the segment coincides with part of a grid line.

@@

\begin{proof}
\ref*{prop: len ineq i j: ineqs} The inequalities follow from the fact that the segment endpoints lie in the interiors or on the outer edges of two tiles in opposite corners of the discrete bounding rectangle. This is illustrated in Figures~\ref{fig: len ineq 4 3} and \ref{fig: len ineq 4 2} for two specific $(i,j)$ pairs respectively. For each $(i,j)$, segments are shown with lengths close to either of the two bounds. Note that inequality \eqref{eq: len > sqrt} is valid even for $i=2$, $j=2$, in which case it reduces to $\len>0$.

\ref*{prop: len ineq i j: exist} For $a$, $b$, $\len$, $i$, $j$ satisfying the two inequalities, a segment of length $\len$ can be found that has its endpoints in the interiors or on the outer edges of the two shaded tiles of a rectangle with normalized dimensions $i$ and $j$, as in Figure~\ref{fig: len ineq i j}, which is thus the discrete bounding rectangle of that segment.
\begin{figure}
\centering%
\subfigure[$i=4$, $j=3$]{%
\label{fig: len ineq 4 3}%
\includegraphics[width=.49\textwidth]{L_ineq_4_3_1p35}%
}\hfill%
\subfigure[$i=4$, $j=2$]{%
\label{fig: len ineq 4 2}%
\includegraphics[width=.49\textwidth]{L_ineq_4_2_1p35}%
}%
\caption{Relationship between segment length $\len$ and dimensions $i$, $j$ of the discrete bounding rectangle
}%
\label{fig: len ineq i j}
\end{figure}%

@@

``There is a segment\ldots $\Rightarrow$ \eqref{eq: len > sqrt} '': Assume that a segment exists with length $\len' \leq \len$ and with a discrete bounding rectangle of normalized dimensions $i$, $j$. From part~\ref{prop: len ineq i j: ineqs}, inequality \eqref{eq: len > sqrt} holds for the length $\len'$, and thus for $\len$.
\end{proof}

Consider the problem of maximizing the number of visited tiles for a given length. According to Proposition~\ref{prop: i+j-1}, the position and orientation of the segment should be chosen to obtain $i+j-1$ as large as possible, where $i$ and $j$ are the normalized dimensions of its discrete bounding rectangle. On the other hand, Proposition~\ref{prop: len ineq i j} restricts the $i, j$ values that can be achieved with a given length. A relevant question is: are there any $(i,j)$ pairs that can be disregarded for all lengths $\len$? In other words, what is the ``smallest'' subset of $\mathbb N^2$ such that the $(i,j)$ pair that maximizes the number of tiles for any given length can always be found within that subset?

@@

Consider $a=b=1$ first. This is illustrated in Figure~\ref{fig: ijLS_1}. Note that in this and in the next figures the axes represent $ia$ and $jb$ (not $i$ and $j$). In this graph, each dashed diagonal line contains points $(ia,jb)$ with the same $i+j-1$; and the lower bound \eqref{eq: len > sqrt} corresponds to an arc centered at $(2a,2b)$.

@@

\begin{proof}
Suppose first that the grid is simplified to only horizontal lines with spacing $b$. This matches the set-up of Buffon's original needle problem \cite{Mathai99}, except that here the length $\len$ of the needle may exceed the spacing $b$, allowing it to cross multiple lines. As shown in \cite{Ramaley69}, the expected number of lines crossed equals $2\len/(\pi b)$.
% When $\len<b$ this expectation matches the probability of the needle crossing a grid line, since it cannot cross more than one.
